A star formation event in the leading arm of the Magellanic stream triggered by passage through the outer Galactic disk

### 2018-09-13 ms.tex
Update on Overleaf.

The halo of the Milky Way is characterized by its old ($\gtrsim 10~\textrm{Gyr}$), metal-poor ($\feh \approx -1.5$) stellar population.
This is understood as a signature of the dominant (in stellar mass) progenitor systems that were accreted and disrupted early on in the formation of the Galaxy: massive dwarf galaxies \citep[e.g.,][]{Deason:2015, Fiorentino:2015}.
These systems likely came in with significant gas reservoirs, but were quenched and stripped through collisional processes that heated and dispersed the gas \citep[e.g.,][]{Mayer:2006}, thus preventing immediate star formation in the deposited gas.
The Milky Way, however, continues to accrete satellite galaxies, as is evidenced by the prominent stellar stream from the Sagittarius dwarf galaxy \citep{Ibata:1994, Majewski:2003}, and the presence of the Large and Small Magellanic Clouds (LMC, SMC) within the virial radius of the Galaxy.
While Sagittarius was likely stripped of its neutral gas long ago \citep{Burton:1999, Tepper-Garcia:2018}, the \lmcsmc\ system is associated with $\approx 8\times 10^8~\msun$ of \hi gas \citep{Bruns:2005}, which extends into leading and trailing gas streams that wrap nearly $\approx 200^\circ$ around the sky \citep{Putman:1998, Bruns:2005, Nidever:2010}.

The Magellanic stream (MS) is a large stream of (predominantly) hydrogen gas emanating from the \lmcsmc\ system that contains a significant fraction of the total gas mass associated with the \lmcsmc\ \citep{Bruns:2005}.
The stream was discovered in early surveys of 21 cm emission and has since been studied in great detail by large-area radio sky surveys, and in H$\alpha$ emission \citep{Weiner:1996, Barger:2017}.
With increased resolution, surveys have found small-scale structure and gas fragmentation \citep[e.g.,][]{Nidever:2008, For:2014} and a large-scale bifurcation, with semi-distinct ``strands'' that lead back to the LMC and broader system, respectively \citep[e.g.,][]{Morras:1983, Putman:2003}.
As surveys push deeper to lower column densities, extensions of the stream have been found both in the trailing region \citep{Nidever:2010}, but also in the leading arm, which has been found to connect to regions of low-column-density gas ($N\sim 10^{18}$--$10^{19}~\textrm{cm}^{-2}$) on the other side of the Galactic disk \citep{Putman:1998, Nidever:2008}.
The leading arm (LA) has been decomposed into three or four distinct ``features'' named LA1--4 \citep{Bruns:2005, Nidever:2008, Venzmer:2012} with total HI mass $\sim 4 \times10^7~\msun$ \citep{Venzmer:2012}.

The origin and formation of the LA features in the MS is still uncertain.
Initial studies of the LA argue that the features closest to the \lmcsmc\ can be traced back to the SMC, thus implying that the gas in the LA was stripped from the SMC \citep{Putman:1998}.
In more recent work, it has instead been argued that outer features of the LMC appear to lead directly into the LA-1 feature \citep{Nidever:2008}, implying an LMC origin for the leading arm.
Whatever the origin of the gas, it is clear that tidal stripping by the Milky Way is required to form the leading arm \citep{Nidever:2008}.
However, the LA features deviate from the predicted orbit of the \lmcsmc, implying that ram pressure or interactions with gas in the outer Milky Way disk may have removed orbital energy from the LA gas \citep[e.g.,][]{Bekki:2008}.

@@

We then query all objects from the \gaia\ \DR{2} catalog within a rectangle centered on the nominal position of the group (see above), with a width of $12^\circ$ and a height of $12^\circ$ in the equatorial (ICRS) coordinate system (see Appendix~\ref{sec:queries} for the database query).
We cross-match the resulting catalog with \acronym{GALEX} \todo{citation} using a positional crossmatch requiring source separation $<2~\textrm{arcsec}$, and with \pans\ \todo{citation}\citep{Chambers:XX} using the cross-match provided by the \gaia\ science archive.
\figurename~\ref{fig:cmds} shows the \gaia\ color-magnitude diagram (left) and the \acronym{GALEX} NUV minus \pans\ $g$-band vs. \gaia\ \bprp\ color-color diagram (right) from the cross-matched photometry.
All magnitudes are corrected for dust extinction using the SFD \citep{Schlegel:1998} dust map.
The green polygon in the left panel of \figurename~\ref{fig:cmds} highlights the over-density of blue stars that were identified as comoving from the proper motion search described above.
The lines in the right panel of \figurename~\ref{fig:cmds} show isochrones from the \acronym{MIST} (blue; \citealt{TODO}) and \acronym{PARSEC} (red; \citealt{TODO}) isochrone libraries for a 50 Myr stellar population with metallicity $\feh = -0.5$ (dashed) or $\feh = -1.5$ (solid).

>>>>

The halo of the Milky Way is characterized by its old ($\gtrsim 10~\textrm{Gyr}$), metal-poor ($\feh \approx -1.5$) stellar population.
This is understood as a signature of the dominant (in stellar mass) progenitor systems that were accreted and disrupted early on in the formation of the Galaxy: massive dwarf galaxies \citep[e.g.,][]{Deason:2015, Fiorentino:2015}.
These systems likely came in with significant gas reservoirs, but were quenched and stripped through collisional processes that heated and dispersed the gas \citep[e.g.,][]{Mayer:2006}, thus preventing immediate star formation in the deposited gas.
The Milky Way, however, continues to accrete satellite galaxies, as is evidenced by the prominent stellar stream from the Sagittarius dwarf galaxy \citep{Ibata:1994, Majewski:2003}, the presence of the Large and Small Magellanic Clouds (LMC, SMC) and about 50 dwarf satellites within the virial radius of the Galaxy.
While Sagittarius was likely stripped of its neutral gas long ago through repeated passage about the Milky Way \citep{Burton:1999, Tepper-Garcia:2018}, the \lmcsmc\ system is associated with $\approx 8\times 10^8~\msun$ of \hi gas \citep{Bruns:2005}, which extends into leading and trailing gas streams that wrap nearly $\approx 200^\circ$ around the sky \citep{Mathewson:1974, Putman:1998, Bruns:2005, Nidever:2010}.

The Magellanic stream (MS) is a large stream of (predominantly) hydrogen gas emanating from the \lmcsmc\ system that contains a significant fraction of the total gas mass associated with the \lmcsmc\ \citep{Bruns:2005}.
The stream was discovered in early surveys of 21 cm emission and has since been studied in great detail by large-area radio sky surveys, and in H$\alpha$ emission \citep{Weiner:1996, Barger:2017}.
With increased resolution, surveys have found small-scale structure and gas fragmentation \citep[e.g.,][]{Nidever:2008, For:2014} and a large-scale bifurcation, with %semi-distinct ``strands'' that lead back to the LMC and broader system, respectively \citep[e.g.,][]{Morras:1983, Putman:2003}.
kinematically \citep{Nidever:2008} and chemically \citep{Fox:2013} distinct ``strands'' that lead back to the LMC and the SMC, respectively.
As surveys push deeper to lower column densities, extensions of the stream have been found both in the trailing region \citep{Nidever:2010}, but also in the leading arm, which has been found to connect to regions of low-column-density gas ($N\sim 10^{18}$--$10^{19}~\textrm{cm}^{-2}$) on the other side of the Galactic disk \citep{Putman:1998, Nidever:2008}.
The leading arm (LA) has been decomposed into three or four distinct ``features'' named LA1--4 \citep{Bruns:2005, Nidever:2008, Venzmer:2012} with total HI mass $\sim 4 \times10^7~\msun$ \citep{Venzmer:2012}.

The origin and formation of the LA features in the MS is still uncertain.
Initial studies of the LA argued that the features closest to the \lmcsmc\ can be traced back to the SMC, thus implying that the gas in the LA was stripped from the SMC \citep{Putman:1998}. 
In later work, it has instead been argued that outer features of the LMC appear to lead directly into the LA-1 feature \citep{Nidever:2008}, implying an LMC origin for the leading arm.
Recent chemical abundance measurements along several sightlines passing through the LA again suggested the SMC origin \citep{Fox:2018,Richter:2018}. 
Hydrodynamical simulations of the MCs on the first infall to the Milky Way, however, showed that the repeated encounters between the MCs strip gas not only from the SMC, but also from the LMC, and this tidally stripped gas creates filamentary gas structures both in the LA and the Magellanic Stream \citep{Pardy:2018}.   
Whatever the origin of the gas, it is clear that tidal stripping by the Milky Way is required to form the leading arm \citep{Nidever:2008, Besla:2012}.
However, the LA features deviate from the predicted orbit of the \lmcsmc, implying that ram pressure or interactions with gas in the outer Milky Way disk may have removed orbital energy from the LA gas \citep[e.g.,][]{Bekki:2008}.

@@

We then query all objects from the \gaia\ \DR{2} catalog within a rectangle centered on the nominal position of the group (see above), with a width of $12^\circ$ and a height of $12^\circ$ in the equatorial (ICRS) coordinate system (see Appendix~\ref{sec:queries} for the database query).
We cross-match the resulting catalog with \acronym{GALEX} \citep{Martin:2005} using a positional cross-match requiring source separation $<2~\textrm{arcsec}$, and with \pans\ \citep{Chambers:2016} using the cross-match provided by the \gaia\ science archive.
\figurename~\ref{fig:cmds} shows the \gaia\ color-magnitude diagram (left) and the \acronym{GALEX} NUV minus \pans\ $g$-band vs. \gaia\ \bprp\ color-color diagram (right) from the cross-matched photometry.
All magnitudes are corrected for dust extinction using the SFD \citep{Schlegel:1998} dust map {\bf (Yumi: Did you use Schlafly's value though?)}.
The green polygon in the left panel of \figurename~\ref{fig:cmds} highlights the over-density of blue stars that were identified as comoving from the proper motion search described above.
The lines in the right panel of \figurename~\ref{fig:cmds} show isochrones from the \acronym{MIST} \citep[blue;][]{Choi:2016} and \acronym{PARSEC} \citep[red;][]{Bressan:2012, Marigo:2017} isochrone libraries for a 50 Myr stellar population with metallicity $\feh = -0.5$ (dashed) or $\feh = -1.5$ (solid).
